On the means, costs, and system-level impacts of 24/7 carbon-free energy procurement

### 2024-01-05 manuscript/manuscript.tex
fix last typos, add citation and corresponding author

	\author[tubaddress]{Iegor Riepin}
	\ead{iegor.riepin@tu-berlin.de}
	\author[tubaddress]{Tom Brown}
	\ead{t.brown@tu-berlin.de}

@@

	\begin{abstract}
		% The 5 things a good abstract needs:
		% 1. Introduce the topic: what we know
		% 2. State the unknown: what we don't know
		% 3. Outline the methods: how we answer the questions
		% 4. Preview the findings: what we learned
		% 5. Tell us what your work teaches us: what we conclude and why this is important
		% aim ~200 words | 225 words used
		A growing number of public and private energy buyers are interested in 24/7 carbon-free energy (CFE) procurement, which means that every kilowatt-hour of electricity consumption is met by carbon-free sources at all times.
		It has the potential to overcome the limitations of established procurement schemes, such as the temporal mismatch between clean electricity supply and buyers' demand that is inherent to \enquote{volumetric} matching.
		Yet it is unclear how 24/7 CFE procurement affects the rest of the power system, and whether this effect is consistent across regional contexts and different levels of system cleanness.
		We use a mathematical model to systematically examine different designs, optimal procurement strategies, costs, and impacts of the 24/7 CFE matching, both for participating buyers and for regions where voluntary procurement occurs.
		We examine mechanisms driving system-level emissions reduction and how they vary across regions and over time.
		Our results indicate that clean energy procurement commitments have consistent beneficial effects on participants and the electricity system.
		Even as grids become cleaner over time, the hourly matching strategy contributes significantly to system-level emissions reduction.
		In addition, voluntary commitments to 24/7 CFE have a further transformative effect on electricity systems through accelerated innovation and early deployment of advanced energy technologies.
	\end{abstract}

@@

Climate change is driving a global effort to rapidly decarbonise electricity systems across the globe.
The European Union (\gls{eu}) has committed to achieving net-zero emissions by 2050, and the European Commission has adopted the European Climate Law with a target of reducing net greenhouse gas emissions by at least 55\% by 2030, compared to 1990 levels \cite{EuropeanClimateLaw2020}.
National climate initiatives are gaining ground across the globe, including the United States with the Biden administration's goal of reducing emissions by 50-52\% by 2030, compared to 2005 levels \cite{BidenClimatePlan2021}, and China's commitment to have CO$_2$ emissions peak before 2030 and achieve carbon neutrality before 2060 \cite{ChinaNetZero-IEA}.

@@

Some corporate and industrial (\gls{ci}) buyers recongnised the problems associated with the unbundled certificates and turned towards Power Purchase Agreements (\gls{ppa}s).
When signing a \gls{ppa}, energy buyers pledge to buy both the energy and the environmental credits \enquote{bundled} to megawatt-hours.
Market-driven procurement of renewable energy has become one of the major drivers of the energy transition: \gls{ppa}s are expected to account for one-fifth of utility solar PV and wind capacity expansion in 2023--2024, and almost twice as much (36\%) when China is excluded \cite{iea-REppa2023}.
In some countries, such as Spain and Sweden, market-driven procurement of renewable energy capacity surpasses policy-driven capacity expansion \cite{iea-REppa2023}.

@@

Google led the 24/7 movement in 2020 by announcing its commitment to cover the demand at all its data centres and campuses worldwide using the 24/7 CFE approach by 2030 \cite{google-247by2030}.
Shortly after, Google published a policy roadmap on achieving the goal \cite{google-PolicyRoadmap}.
Also, Microsoft has committed to the 100/100/0 by 2030 goal (i.e. \enquote{100\% of electricity consumption, 100\% of the time, matched by zero carbon energy purchases} \cite{Microsoft-vision}.
The list of entities that have committed to 24/7 CFE procurement includes also utilities, such as Peninsula Clean Energy (a community choice aggregator in California) committed to 24/7~CFE by 2025 \cite{peninsula-OurPathto247}, and even cities like Des Moines, Iowa U.S., with a commitment by 2035 \cite{iowaenvcouncil-247}. In 2021, an international group of energy suppliers, electricity buyers, and \gls{ngo}s launched the 24/7 Carbon-free Energy Compact \cite{gocarbonfree247}.
With over 135 signatories, the group aims to \enquote{develop and scale high-impact technologies, energy policies, procurement practices, aand solutions to make 24/7 Carbon-Free Energy achievable for all}.

@@

% On bilinear term due to grid CFE factor
The CFE factor of the regional grid ($CFE_t$) can be seen as the percentage of clean electricity in each MWh of imported electricity to supply demand of participating consumers in a given hour.
To compute $CFE_t$, we consider both the hourly electricity mix in the local bidding zone and the emissionality of imported electricity from the neighbouring zones.
The methodology to calculate the grid CFE factor is described in the prior work of the authors \cite{riepin-zenodo-systemlevel247}.

@@

\textbf{Technology assumptions --} Cost and other assumptions for energy technologies available for participating consumers were collected primarily from the Danish Energy Agency \cite{DEA-technologydata} for the respective years and are provided in [\cref{tab:tech_costs}].
For the \gls{ldes}, we assume an underground hydrogen storage in salt caverns with an electolyzer and a fuel cell for hydrogen conversion.
Data for advanced clean firm technologies is less reliable due to technological uncertainty and lack of commercial experience; therefore, use our own indicative assumptions.

@@

    \caption{
        System-level emission reduction: comparison of procurement policies and isolation of decarbonisation mechanisms.
        \cref{fig:10-profile-volume.pdf} shows breakdown of emissions reduction in a local zone if 10\% of \gls{ci} load follows 24/7 hourly procurement with CFE 100\% target. The profile are volume mechanisms are isolated.
        \cref{fig:10-hourly-annual.pdf} compares reductions in annual zone emissions in percent points achieved through 100\% annual renewable matching and 100\% 24/7 CFE matching; no procurement policy is assumed in the counterfactual scenario.}
        \label{fig:decarbonisation_story}

@@

\textbf{Our results in perspective --} Our findings agree with the existing model-based studies on clean energy procurement \cite{xu-247CFE-SSRN,ieaAdvancingDecarbonisationClean2022,peninsula-report247}, as well as with the general understanding of the energy research community that \enquote{100\% renewable energy is not enough} \cite{chalendar-2019} in a context of electricity consumers claiming environmental sustainability.
Specifically, we show that through hourly matching of demand with clean electricity, electricity consumers can both completely negate their carbon emissions and contribute to broader system-wide decarbonisation.
100\% CFE matching is viable with commercially available technologies, albeit incurs a cost premium for participating consumers.
The cost premium can be substantially reduced if the advanced technologies, such as long-duration energy storage and clean firm generators are available, or if consumers opt for participating in the 24/7~CFE procurement with quality scores ranging from 90\% to 95\% instead of an absolute 100\%.

@@

% Limitations: consumption profiles
Third, we assume that the participating consumers have a flat demand profile (i.e. baseload).
As a matter of fact, consumer demand profiles vary based on their business activities.
It has been shown by Riepin~\&~Brown (2022) that the shape of consumption profiles affects the cost-optimal technology mix required for achieving a certain CFE target; however, the shape of the profile has relatively little impact on procurement costs, emissionality of the portfolio, and system-level impacts of 24/7 procurement \cite{riepin-zenodo-systemlevel247}.

@@

% Excess constraint problem
One of the simplifications of mathematical models implementing 24/7 CFE procurement is how the excess carbon-free electricity is handled.
The \enquote{excess CFE} can be defined as the total amount of carbon-free electricity supply exceeding the demand of the participating consumers in a given period of time.
Here we assume that the excess can either be curtailed or sold to the regional electricity market at wholesale market prices, until the excess constraint is reached.
The constraint restricts the volume of excess electricity sold to the regional grid, ensuring it does not surpass 20\% of the annual demand from participating consumers.
The problem with the uncostrained export of excess CFE to the regional market could be a situation where excess electricity from the renewable generators procured by 24/7 consumers displaces electricity from other renewable generators in the market, which in turn may challenge additionality of the 24/7 procurement; and if system-wide renewable expansion is constrained, the cost-optimal solution may be to use the excess clean electricity to \enquote{relax} the constraint, increasing the renewable portfolio of participating consumers.
Similar \enquote{spillover effect} problem has been observed and discussed by Xu et al. (2022) \cite{xu-247CFE-SSRN}.
In their white paper, Peninsula Clean Energy team also discuss the problems and implications of the excess supply. Citing their work: \enquote{We conclude that excess supply is a necessary aspect of a time-coincident renewable portfolio in the range of likely market conditions in California today} \cite{peninsula-report247}.
Both studies perform sensitivity analyses on the excess supply and discuss the implications on their findings \cite{xu-247CFE-SSRN, peninsula-report247}.

@@

Another alternative strategy could be that companies do annual matching with \enquote{overprocurement}, i.e., covering their own consumption on an annual volumentric basis plus procuring some more clean electricity on top. Such strategy could be implemented at low cost, e.g., with solar PV.
However, as shown in \cref{subsec:mechanisms}, this strategy would not be as effective in reducing system-level emissions as 24/7 CFE matching, especially in the context of evolving national energy and climate policies.
Further, such a strategy would not promote innovation and pull forward in time research and investment decisions in advanced energy technologies, such as long-duration energy storage and clean firm generators.

>>>>

	\author[tubaddress]{Iegor Riepin\fnref{fn1}}
	\ead{iegor.riepin@tu-berlin.de}

	\author[tubaddress]{Tom Brown}
	\ead{t.brown@tu-berlin.de}

@@

    \fntext[fn1]{Corresponding author: Iegor Riepin, Technical University of Berlin, Einsteinufer 25, 10587 Berlin, Germany}

	\begin{abstract}
		% The 5 things a good abstract needs:
		% 1. Introduce the topic: what we know
		% 2. State the unknown: what we don't know
		% 3. Outline the methods: how we answer the questions
		% 4. Preview the findings: what we learned
		% 5. Tell us what your work teaches us: what we conclude and why this is important
		% aim ~200 words | 225 words used
		A growing number of public and private energy buyers are interested in 24/7 carbon-free energy (CFE) procurement, which means that every kilowatt-hour of electricity consumption is met by carbon-free sources at all times.
		It has the potential to overcome the limitations of established procurement schemes, such as the temporal mismatch between clean electricity supply and buyers' demand that is inherent to \enquote{volumetric} matching.
		Yet it is unclear how 24/7 CFE procurement affects the rest of the power system, and whether this effect is consistent across regional contexts and different levels of system cleanness.
		We use a mathematical model to systematically examine different designs, optimal procurement strategies, costs, and impacts of the 24/7 CFE matching, both for participating buyers and for regions where voluntary procurement occurs.
		We examine mechanisms driving system-level emissions reduction and how they vary across regions and over time.
		Our results indicate that clean energy procurement commitments have consistent beneficial effects on participants and the electricity system.
		Even as grids become cleaner over time, the hourly matching strategy contributes significantly to system-level emissions reduction.
		In addition, voluntary commitments to 24/7 CFE have a further transformative effect on electricity systems through accelerated innovation and early deployment of advanced energy technologies.

	\end{abstract}

@@

Climate change is driving a global effort to rapidly decarbonise electricity systems across the globe.
The European Union (\gls{eu}) has committed to achieving net-zero emissions by 2050, and the European Commission has adopted the European Climate Law with a target of reducing net greenhouse gas emissions by at least 55\% by 2030, compared to 1990 levels \cite{EuropeanClimateLaw2020}.
National climate initiatives are gaining ground across the globe, including the United States with the Biden administration's goal of reducing emissions by 50-52\% by 2030, compared to 2005 levels \cite{BidenClimatePlan2021}, and China's commitment to have CO$_2$ emissions peak before 2030 and achieve carbon neutrality before 2060 \cite{ChinaNetZero-IEA, liuEconomicCostEnergy2023}.

@@

Some corporate and industrial (\gls{ci}) buyers recognised the problems associated with the unbundled certificates and turned towards Power Purchase Agreements (\gls{ppa}s).
When signing a \gls{ppa}, energy buyers pledge to buy both the energy and the environmental credits \enquote{bundled} to megawatt-hours.
Market-driven procurement of renewable energy has become one of the major drivers of the energy transition: \gls{ppa}s are expected to account for one-fifth of utility solar PV and wind capacity expansion in 2023--2024, and almost twice as much (36\%) when China is excluded \cite{iea-REppa2023}.
In some countries, such as Spain and Sweden, market-driven procurement of renewable energy capacity surpasses policy-driven capacity expansion \cite{iea-REppa2023}.

@@

Google initiated the 24/7 movement in 2020 by announcing its commitment to cover the demand at all its data centres and campuses worldwide using the 24/7 CFE approach by 2030 \cite{google-247by2030}.
Shortly after, Google published a policy roadmap on achieving the goal \cite{google-PolicyRoadmap} and methodologies paper for how to measure the progress \cite{google-methodologies}.
Also, Microsoft has committed to the 100/100/0 by 2030 goal (i.e. \enquote{100\% of electricity consumption, 100\% of the time, matched by zero carbon energy purchases} \cite{Microsoft-vision}.
The list of entities that have committed to 24/7 CFE procurement includes also utilities, such as Peninsula Clean Energy (a community choice aggregator in California) committed to 24/7~CFE by 2025 \cite{peninsula-OurPathto247}, and even cities like Des Moines, Iowa U.S., with a commitment by 2035 \cite{iowaenvcouncil-247}. In 2021, an international group of energy suppliers, electricity buyers, and \gls{ngo}s launched the 24/7 Carbon-free Energy Compact \cite{gocarbonfree247}.
With over 135 signatories, the group aims to \enquote{develop and scale high-impact technologies, energy policies, procurement practices, and solutions to make 24/7 Carbon-Free Energy achievable for all}.

@@

% On bilinear term due to grid CFE factor
The CFE factor of the regional grid ($CFE_t$) can be seen as the percentage of clean electricity in each MWh of imported electricity to supply demand of participating consumers in a given hour.
To compute $CFE_t$, we consider both the hourly electricity mix in the local bidding zone and emission intensity of imported electricity from the neighbouring zones.
The methodology to calculate the CFE factor of the regional grid is described in detail in the prior work of the authors \cite{riepin-zenodo-systemlevel247}.

@@

\textbf{Technology assumptions --} Cost and other assumptions for energy technologies available for participating consumers were collected primarily from the Danish Energy Agency \cite{DEA-technologydata} for the respective years and are provided in [\cref{tab:tech_costs}].
For the \gls{ldes}, we assume an underground hydrogen storage in salt caverns with an electolyzer and a fuel cell for hydrogen conversion.
Data for advanced clean firm technologies is less reliable due to technological uncertainty and lack of commercial experience; therefore, we use our own indicative assumptions.

@@

    \caption{
        System-level emission reduction: comparison of procurement policies and isolation of decarbonisation mechanisms.
        \cref{fig:10-profile-volume.pdf} shows breakdown of emissions reduction in a local zone if 10\% of \gls{ci} load follows 24/7 hourly procurement with CFE 100\% target. The profile and volume mechanisms are isolated.
        \cref{fig:10-hourly-annual.pdf} compares reductions in annual zone emissions in percent points achieved through 100\% annual renewable matching and 100\% 24/7 CFE matching; no procurement policy is assumed in the counterfactual scenario.}
        \label{fig:decarbonisation_story}

@@

\textbf{Our results in perspective --} Our findings agree with the existing model-based studies on clean energy procurement \cite{xu-247CFE-SSRN,ieaAdvancingDecarbonisationClean2022,peninsula-report247}, as well as with the general understanding of the energy research community that \enquote{100\% renewable energy is not enough} \cite{chalendar-2019} in a context of electricity consumers claiming complete environmental sustainability.
Specifically, we show that through hourly matching of demand with clean electricity, electricity consumers can both completely negate their carbon emissions and contribute to broader system-wide decarbonisation.
100\% CFE matching is viable with commercially available technologies, albeit incurs a cost premium for participating consumers.
The cost premium can be substantially reduced if the advanced technologies, such as long-duration energy storage and clean firm generators are available, or if consumers opt for participating in the 24/7~CFE procurement with quality scores ranging from 90\% to 95\% instead of an absolute 100\%.

@@

% Limitations: consumption profiles
Third, we assume that the participating consumers have a flat demand profile (i.e. baseload).
As a matter of fact, consumer demand profiles vary based on their business activities.
It has been shown by Riepin~\&~Brown (2022) that the shape of consumption profiles affects the cost-optimal technology mix required for achieving a certain CFE target; however, the shape of the profile has relatively little impact on procurement costs, cost-optimal portfolio of technologies, and system-level impacts of 24/7 procurement \cite{riepin-zenodo-systemlevel247}.

@@

% Excess constraint problem
One of the simplifications of mathematical models implementing 24/7 CFE procurement is how the excess carbon-free electricity is handled.
The \enquote{excess CFE} can be defined as the total amount of carbon-free electricity supply exceeding the demand of the participating consumers in a given period of time.
Here we assume that the excess can either be curtailed or sold to the regional electricity market at wholesale market prices, until the excess constraint is reached.
The constraint restricts the volume of excess electricity sold to the regional grid, ensuring it does not surpass 20\% of the annual demand from participating consumers.
The problem with the unconstrained export of excess CFE to the regional market could be a situation where excess electricity from the renewable generators procured by 24/7 consumers displaces electricity from other renewable generators in the market, which in turn may challenge additionality of the 24/7 procurement; and if system-wide renewable expansion is constrained, the cost-optimal solution may be to use the excess clean electricity to \enquote{relax} the constraint, increasing the renewable portfolio of participating consumers.
Similar \enquote{spillover effect} problem has been observed and discussed by Xu et al. (2022) \cite{xu-247CFE-SSRN}.
In their white paper, Peninsula Clean Energy team also discuss the problems and implications of the excess supply. Citing their work: \enquote{We conclude that excess supply is a necessary aspect of a time-coincident renewable portfolio in the range of likely market conditions in California today} \cite{peninsula-report247}.
Both studies perform sensitivity analyses on the excess supply and discuss the implications on their findings \cite{xu-247CFE-SSRN, peninsula-report247}.

@@

Another alternative strategy could be that companies do annual matching with \enquote{overprocurement}, i.e., covering their own consumption on an annual volumetric basis plus procuring some more clean electricity on top. Such strategy could be implemented at low cost, e.g., with solar PV.
However, as shown in \cref{subsec:mechanisms}, this strategy would not be as effective in reducing system-level emissions as 24/7 CFE matching, especially in the context of evolving national energy and climate policies.
Further, such a strategy would not promote innovation and pull forward in time research and investment decisions in advanced energy technologies, such as long-duration energy storage and clean firm generators.
